\mytitle

### 2016-04-26 report/paper.tex
Fix some grammar errors

Ordering of events is one the fundamental problems of distributed systems. Creating an algorithm which provide scalability, integrity and validity, along with an ordering for the event through all the peers and processes is one the hot topics in distributed systems research areas for many years. One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}. EpTO is an algorithm which is claimed to provide integrity, validity and total order throughout a distributed system. In addition, EpTO is designed to work without a global clock or synchronization process, that is why EpTO is well-suited for dynamic large-scale distributed systems.

@@

There are many other algorithm for dissemination and ordering events in a distributed systems. There are some deterministic algorithms, which guarantee total order, agreement or other strong properties deterministically. But this type of algorithms are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.

@@

As an alternative for deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}. These algorithms guarantee to disseminate events through the system with high probability. In this way, there is not need for failure detectors and redundant traffic. Thus, these algorithms can be quit scalable. Although because these type of algorithms focus on reliability of dissemination, they often ignore other needed properties such as ordering.

@@

This is where EpTO is coming into the light. EpTO with mixing these two types of algorithms, provide total order along with scalability, validity and integrity. EpTO consists of two parts. First part is probabilistic dissemination. EpTO guarantees that all peers will receive an event with high probability. Second part of EpTO is deterministic ordering. Once peers receive all events, they will deterministically order them using the events timestamp, and in case of a tie using event's broadcaster's id.

@@

For the first part, EpTO is using \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}. The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin completely random and independent from other balls. In this scenario, one the natural questions that comes to mind is that: what is the minimum number of balls that should be thrown, so that every bin has at least one ball?

@@

Using the balls-and-bins approach we model processes as bins and events as balls. Based on this model we will calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability. Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes. Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provides total order.

\subsection{Contribution}
Until now, the creators of EpTO test this algorithm in a simulated environment, and not in a real-world scenario. We are going implement EpTO in pure JAVA using NeEM library \autocite{neem}. Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems. This comparison help us to verify if EpTO is actually perform as expected in real-world conditions or not.
\par
These tasks will be split into two parts. The first part will assume that all peers are known by all peers. We will implement EpTO using this assumption and test it against JGroups. I the second part of the tasks, will assume unknown membership on all peers. We will update EpTO and use a PSS\footnote{Peer Sampling Service}. We will also update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.

>>>>

The ordering of events is one the most prominent problems in distributed systems \todo[inline]{cite paper}. Creating an algorithm which provides scalability, integrity and validity, along with an ordering for the events through all peers and processes has been one the hot topics in distributed systems research for many years. One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}. EpTO is an algorithm which claims to provide integrity, validity and total order in a large scale distributed system. In addition, EpTO is designed to work without a global clock or synchronization process, that is why EpTO is well-suited for dynamic large-scale distributed systems.

@@

There are many other algorithms for disseminating and ordering events in a distributed system. There are some deterministic algorithms, which guarantee total order, agreement or other strong properties. Unfortunately, these types of algorithm are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.

@@

As an alternative for deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}. These algorithms guarantee the dissemination of events in the system with high probability. This way, there is no need for failure detectors and redundant traffic thus, these algorithms are highly scalable. Although because these type of algorithms focus on reliability of dissemination, they often ignore other needed properties such as total ordering.

@@

This is where EpTO comes into light. EpTO, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
EpTO consists of two parts. The first part is probabilistic dissemination. EpTO guarantees that all peers will receive an event with an arbitrarily high probability. The second part of EpTO is deterministic ordering. Once peers have received all events, they will deterministically order them using the events timestamp, and in case of a tie use the event's broadcaster id.

@@

For the first part, EpTO is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}. The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin completely random and independent from other balls. In this scenario, one the natural questions that comes to mind is that: what is the minimum number of balls that should be thrown, so that every bin has at least one ball?

@@

Using the balls-and-bins approach we model processes as bins and events as balls. Based on this model we will calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability. Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes. Using these approaches, EpTO becomes highly scalable and resilient, while still providing total order.

\subsection{Contribution}
Until now, the creators of EpTO have only tested this algorithm in a simulated environment. We are going to implement EpTO in pure JAVA using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems. Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems. This comparison helps us verify if EpTO is actually performing as expected in a real-world scenario.
\par
These tasks will be split in two parts. The first part will assume that all peers are known by all peers. We will implement EpTO using this assumption and test it against JGroups. In the second part of the tasks, we will assume unknown membership on all peers. We will update EpTO and use a PSS\footnote{Peer Sampling Service}. We will also update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
